### 2023-03-21 Cap3.tex

\section{OPR Circuits}
\begin{itemize}
\item How can we reduce the amount of parametrized rotations used?
\item \textbf{OPR Circuit:} Define an OPR circuit (a circuit with one parametrized rotation)
\item \textbf{Theorem about OPR circuits:} The theorem gives the general form that the operator of an OPR circuit can have. \textit{An OPR circuit on $n$ qubits with parameter $p$ always has an operator of the form:}
\begin{align*}
U = (\vec{v}_0, \vec{v}_1, \cdots, \vec{v}_{2^n-1})
\end{align*}
\textit{where $s = s(p)$ and the column vectors are $\vec{v}_j = e^{is} \vec{a}_j + e^{-is} \vec{b}_j + \vec{c}_j$, with $\vec{a}_j, \vec{b}_j, \vec{c}_j$ orthogonal vectors and $|\vec{a}_j|^2 + |\vec{b}_j|^2 + |\vec{c}_j|^2 = 1$.}
\end{itemize}