### 2023-03-21 Cap5.tex

\section{Exponential maps $e^{iHt}$}
\begin{itemize}
\item We will apply the results about OPR circuit to another kind of operation.
\item Consider operations $U  = e^{iHt}$ where $H$ is hermitian  and time independent.
\item \textbf{Theorem:} A n-qubit operator of the form $U = e^{iHt}$ can be implemented using an OPR circuit if and only if the eigenvalues of $H$ are $-\lambda,0,\lambda$ (for $\lambda$ a real number) and the degeneracies of $\lambda$ and $-\lambda$ are $2^j$ for some $j\in \{0,1,\cdots,n-1\}$.
\item \textbf{Examples:} Every matrix $H$ for one qubit, any $n$-qubit pauli matrix, some sums of Pauli matrices, etc.
\end{itemize}